---
title: "R Notebook"
output: html_notebook
---

Introduction of package ClusterR.
Documentation can be found at https://cran.r-project.org/web/packages/ClusterR/vignettes/the_clusterR_package.html.


---> dataset sintetici 2d per visualizzazioni 
  punti generati da gaussiane
  punti generati da PCA di un dataset grande -> proteine? 
  
  
  

-  Introduzione al clustering (cos'è, immagini d'esempio, esempi di applicazione)
-  Descrizione dei tipi di clustering 
-       Connectivity based
-       Centroid based
-       Distribution based
-       Density based
-  Dataset di esempio della libreria
-  Clustering supportati da ClusterR (centroid based, distribution based)
-  Centroid based clustering (k-means, mini-batch-kmeans, k-medoids)
-       k-means basic usage - KMeans_rcpp
-       variare il numero di k - immagini per diversi k
-       optimal k - Optimal_Clusters_KMeans 
-       visualizzare la bonta del clustering - Silouhette plot
-       misura del tempo di esecuzione - large dataset - KMeans_rcpp vs KMeans_arma vs mini-batck KMeans
-       medoids vs centroids - Cluster_Medoids
-       misura del tempo di esecuzione - Cluster_Medoids vs Clara_Medoids
-  Distribution based clustering
-       Gaussian Mixture Models Optimal_Clusters_GMM 
-       visualizzazione 2d gaussiane
-  Distribution vs Centroid
-       differenze clustering usando i dataset sintetici
-       external_validation per validazione con true label 
-  (optional) paragone ClusteR vs ClusterR




```{r}


# load library MASS
library(MASS)
library(dplyr)
library(ClusterR)

load_factor = 100
ncol=2

coeff = log(base=10, load_factor)/2    #choose if augmenting distance when generating more points
#coeff = 1
  
# create bivariate normal distribution
df1 = bind_rows(list(
  data.frame(mvrnorm(n = 10 * load_factor, mu = c(5*coeff, 5*coeff), Sigma = matrix(c(1, -.6, 0, 1), ncol = 2)), color=1),
  data.frame(mvrnorm(n = 10 * load_factor, mu = c(0*coeff, 0*coeff), Sigma = matrix(c(.8, 0, 0, 4), ncol = 2)), color=2),
  data.frame(mvrnorm(n = 15 * load_factor, mu = c(3*coeff, 1*coeff), Sigma = matrix(c(3, .7, 0, 2), ncol = 2)), color=3)
))

plot(x=df1$X1, y=df1$X2, col=df1$color)
points(c(5*coeff), c(5*coeff), pch="1", cex=2, col="brown")
points(c(0*coeff), c(0*coeff), pch="2", cex=2, col="brown")
points(c(3*coeff), c(1*coeff), pch="3", cex=2, col="brown")

```







```{r}

x = subset(df1, select=-c(color))
gmm_3 = GMM(x, 3, dist_mode = "maha_dist", seed_mode = "random_subset")          

# predict centroids, covariance matrix and weights
df1["color_pred_3"] = predict(gmm_3, newdata = x)

plot(x=df1$X1, y=df1$X2, col=df1$color_pred_3 + 1)
points(c(5*coeff), c(5*coeff), pch="1", cex=2, col="black")
points(c(0*coeff), c(0*coeff), pch="2", cex=2, col="black")
points(c(3*coeff), c(1*coeff), pch="3", cex=2, col="black")
points(gmm_3$centroids, pch="X", cex=2, col="azure")

```


```{r}
plot_2d(data = x, clusters = df1$color_pred_3, centroids_medoids = gmm_3$centroids)
```





```{r}
external_validation(df1$color, df1$color_pred_3, summary_stats = T)
external_validation(df1$color, df1$color_pred_2, summary_stats = T)
```




```{r}

x = subset(df, select=-c(color))
gmm_2 = GMM(x, 2, dist_mode = "maha_dist", seed_mode = "random_subset")          

# predict centroids, covariance matrix and weights
df1["color_pred_2"] = predict(gmm_2, newdata = x)

plot(x=df1$X1, y=df1$X2, col=df1$color_pred_2 + 1)
points(c(5*coeff), c(5*coeff), pch="1", cex=2, col="black")
points(c(0*coeff), c(0*coeff), pch="2", cex=2, col="black")
points(c(3*coeff), c(1*coeff), pch="3", cex=2, col="black")
points(gmm_2$centroids, pch="X", cex=2, col="azure")

```

```{r}
plot_2d(data = x, clusters = df1$color_pred_2, centroids_medoids = gmm_2$centroids)

help(plot_2d)
```




```{r}
gmm_3$centroids
gmm_3$covariance_matrices
head(gmm_3$Log_likelihood)
```



```{r}

x = subset(df, select=-c(color))
opt_gmm = Optimal_Clusters_GMM(x, max_clusters = 10, criterion = "BIC", 
                               dist_mode = "eucl_dist", seed_mode = "random_subset",
                               plot_data = T)      


help(Optimal_Clusters_GMM)

```





```{r}

help(Silhouette_Dissimilarity_Plot)

data(soybean)

dat = soybean[, -ncol(soybean)]

cm = Cluster_Medoids(dat, clusters = 5, distance_metric = 'jaccard_coefficient')

plt_sd = Silhouette_Dissimilarity_Plot(cm, silhouette = TRUE)
```




In the package ClusterR there are 3 different datasets: 
-   dietary_survey_IBS: Synthetic data using a dietary survey of patients with irritable bowel syndrome. 
-   mushroom: The mushroom data
-   soybean: The soybean (large) data set from the UCI repository


*dietary_survey_IBS* contains synthetic data generated using the mean and standard deviation of data used in the paper "A dietary survey of patients with irritable bowel syndrome".

```{r}
help(dietary_survey_IBS)
```

It contains 400 rows and 43 columns. 

"class" is the target column. Its values are O and 1, with 0 meaning healty and 1 meaning IBS-positive. 

The other 42 columns are numeric and can be used as predictors. 


```{r}
dim(dietary_survey_IBS)

colnames(dietary_survey_IBS)

unique(dietary_survey_IBS$class)

head(dietary_survey_IBS)
```

In the following a 2d PCA for the entire dataset: black for class 0 (healty) and red for class 1 (positive). 

```{r}

pca_dat = stats::princomp(dietary_survey_IBS)$scores[, 1:2]
plot(pca_dat, col=dietary_survey_IBS$class + 1)

```





```{r}
help(dietary_survey_IBS)
help(mushroom)
help(soybean)
```





*mushroom* includes descriptions of hypothetical samples corresponding to 23 species of gilled mushrooms in the Agaricus and Lepiota Family (pp. 500-525). 

Each species can either be edible or poisonous. The Guide clearly states that there is no simple rule for determining the edibility of a mushroom; no rule like 'leaflets three, let it be' for Poisonous Oak and Ivy. For simplicity the unknown targets are marked as poisounous. 


```{r}
help(mushroom)
```

It contains 8124 rows and 23 columns. 

"class" is the target column. Its values are "p" and "e", with "p" meaning poisounous (or unknown) and e meaning "edible". 

The other 22 columns contain one character each and can be used as predictors. 


```{r}
dim(mushroom)

colnames(mushroom)

unique(mushroom$class)

head(mushroom)
```

In the following a 2d PCA for the entire dataset: black for class 0 (edible) and red for class 1 (poisounous). 

```{r}
X = mushroom[1:1000, -1]

y = as.numeric(mushroom[1:1000, 1])            

gwd = FD::gowdis(X)           

gwd_mat = as.matrix(gwd)                 

pca_dat = stats::princomp(gwd_mat)$scores[, 1:2]
plot(pca_dat, col=y)

```







TODO soybean dataset exploration





















